---
title: "Confirmed Cases and Active Cases by Province"
output: html_notebook
---

```{r Setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
library(utils)
library(lubridate)
library(ggplot2)
library(xts)
library(timeDate)
```

#Confirmed Cases Data
```{r}
confirmed_cases_data <- read.csv("https://raw.githubusercontent.com/dsfsi/covid19za/master/data/covid19za_provincial_cumulative_timeline_confirmed.csv", na.strings = "", fileEncoding = "UTF-8-BOM", stringsAsFactors = TRUE)
confirmed_cases_data
```

#Write raw data
```{r}
write.csv(confirmed_cases_data, "data/data_raw/raw_08_confirmed_and_active_cases_by_province/raw_cases.csv")
```

#Pivot longer
```{r}
confirmed_cases_data <- confirmed_cases_data %>% 
  pivot_longer(c('EC', 'FS','GP', 'KZN', 'LP', 'MP', 'NC', 'NW', 'WC'), names_to = "province", values_to = "confirmed_cases")
confirmed_cases_data
```

#Rename columns and convert date to date format
```{r}
confirmed_cases_data <- confirmed_cases_data %>%
  mutate(date = dmy(date),
         total_cases = total) %>% 
  select(date, province, confirmed_cases)
confirmed_cases_data
```

#Recoveries Data
```{r}
recoveries_data <- read.csv("https://raw.githubusercontent.com/dsfsi/covid19za/master/data/covid19za_provincial_cumulative_timeline_recoveries.csv", na.strings = "", fileEncoding = "UTF-8-BOM", stringsAsFactors = TRUE)
recoveries_data
```

#Write raw data
```{r}
write.csv(recoveries_data, "data/data_raw/raw_08_confirmed_and_active_cases_by_province/raw_recoveries.csv")
```

#Pivot longer
```{r}
recoveries_data <- recoveries_data %>% 
  pivot_longer(c('EC', 'FS','GP', 'KZN', 'LP', 'MP', 'NC', 'NW', 'WC',), names_to = "province", values_to = "recoveries")
recoveries_data
```

#Rename columns and convert date to date format
```{r}
recoveries_data <- recoveries_data %>%
  mutate(date = dmy(date),
         total_recoveries = total) %>% 
  select(date, province, recoveries) 
recoveries_data
```

#Deaths Data
```{r}
deaths_data <- read.csv("https://raw.githubusercontent.com/dsfsi/covid19za/master/data/covid19za_provincial_cumulative_timeline_deaths.csv", na.strings = "", fileEncoding = "UTF-8-BOM", stringsAsFactors = TRUE)
deaths_data
```

#Write raw data
```{r}
write.csv(deaths_data, "data/data_raw/raw_08_confirmed_and_active_cases_by_province/raw_deaths.csv")
```

#Pivot longer
```{r}
deaths_data <- deaths_data %>% 
  pivot_longer(c('EC', 'FS','GP', 'KZN', 'LP', 'MP', 'NC', 'NW', 'WC',), names_to = "province", values_to = "deaths")
deaths_data
```

#Rename columns and convert date to date format
```{r}
deaths_data <- deaths_data %>%
  mutate(date = dmy(date),
         total_deaths = total) %>% 
  select(date, province, deaths)
deaths_data
```

#Join recoveries data to confirmed cases data
```{r}
active_cases_data <- recoveries_data %>% 
  left_join(confirmed_cases_data)
active_cases_data
```
#Join previous table to deaths data 
```{r}
active_cases_data <- active_cases_data %>% 
  left_join(deaths_data) %>% 
  mutate(active_cases = confirmed_cases - (recoveries + deaths)) %>% 
  select(date,province, confirmed_cases, active_cases)
active_cases_data
```

#Rename data
```{r}
active_cases_data <- active_cases_data %>% 
  mutate(province = str_replace(province, "WC", "WESTERN CAPE"),
         province = str_replace(province, "EC", "EASTERN CAPE"),
         province = str_replace(province, "GP", "GAUTENG"),
         province = str_replace(province, "KZN", "KWAZULU NATAL"),
         province = str_replace(province, "NW", "NORTH WEST"),
         province = str_replace(province, "FS", "FREE STATE"),
         province = str_replace(province, "NC", "NORTHERN CAPE"),
         province = str_replace(province, "MP", "MPUMALANGA"),
         province = str_replace(province, "LP", "LIMPOPO"))
active_cases_data
```

```{r}

```



#Write to data_out folder
```{r}
saveRDS(active_cases_data, "data/data_out/08_confirmed_cases_and_active_cases_by_province.rds")
write.csv(active_cases_data, "data/data_out/08_confirmed_cases_and_active_cases_by_province.csv")
```

<!-- Incomplete -->

```{r}
plot <- ggplot() + 
  geom_line(data = active_cases_data,
            aes(x = date,
                y = confirmed_cases),
            color = "#b54c49") +
  geom_line(data = active_cases_data,
            aes(x= date,
                y = active_cases),
            color = "#3484c4") +
#  geom_line(aes(y = active_cases))# +
#  geom_text(aes(label = confirmed_cases), 
#            position=position_dodge(width=0.9), 
#            vjust=-0.25,
#           color = "white") #+
  labs(x = NULL,
       y = NULL,
       title = "Confirmed Cases vs Active Cases by Province",
       caption = "Updated: 16 June 2020") +
  scale_y_continuous(limits = c(0,40000), breaks = seq(0,38000, 19000)) +
  theme_minimal() + 
  theme(#axis.text.y = element_text(color = "white", size = 7),
        axis.text.y = element_blank(),
        axis.text.x = element_text(color = "white"),
        axis.text.x.bottom = element_blank(),
        plot.background = element_rect(fill = "#1c344c", color = "#142c44", size = 5),
        plot.title = element_text(face = "bold", 
                                  size =rel(1.5), 
                                  color = "white"),
        plot.caption = element_text(size = 9, hjust = 0, face = "bold", color = "white"),
        panel.grid = element_blank(),
#        panel.grid.major.y = element_line(linetype = "dotted",
#                                          color = "lightblue"),
        panel.background = element_rect(fill="#1c344c", 
                                        color = "#1c344c"),
        strip.background = element_rect(fill="#1c344c", 
                                        color = "#1c344c"),
        strip.text = element_text(colour = "white",
                                  face = "bold"))+
  facet_wrap(~province, scales = "free_y")
plot
```
















